### 2018-06-04 Writing/draft_v1.tex
draftv1 backup

\begin{figure}
\centering
   \includegraphics[width=0.9\textwidth]{outline.eps}
\caption{Outline of the simulation protocol: the first step includes both the simulation and the modification of the matrices (thin solid lines); the second step includes tree inference using maximum parsimony and Bayesian inference methods (thick solid lines); the third step includes comparing the resulting tree topologies (dashed lines). $n$ and $m$ corresponds to the number of characters with a character difference $<0.25$ and $>0.75$ respectively.}
\label{Fig:outline}
\end{figure}

@@

\begin{figure}
\centering
   \includegraphics[width=0.9\textwidth]{Modif_matrix.eps}
\caption{Example illustration of the protocol for modifying matrices. The matrices represent the pairwise character differences for 100 characters. Blue colours correspond to low character differences and orange colours correspond to high character differences. \textbf{A} - a random Birth-Death tree is simulated and used for generating the ``normal'' matrix (\textbf{B}), characters in this matrix are then removed or duplicated to favour maximised (\textbf{C}), minimised (\textbf{D}) or randomise character difference (\textbf{E}). The differences between the characters is low in \textbf{C} (minimised compared to \textbf{A}) implying a high correlation between the characters. Conversely, the character differences is high in \textbf{D} (maximised compared to \textbf{A}) implying a low correlation between the characters.}
\label{Fig:modif_matrix}
\end{figure}

\subsection{Inferring topologies}
We inferred the topologies with both Bayesian inference and maximum parsimony using MrBayes \citep[v3.2.6;][]{Ronquist2012mrbayes} and PAUP* \citep[v4.0a151;][]{swofford2001paup} respectively.
For both methods, we used the arbitrarily chosen outgroup in the simulations to root our trees.
The maximum parsimony inference was run using a heuristic search with random sequence addition replicate 100 times with a limit of $5\times10^6$ rearrangements per replicates (\texttt{hsearch addseq=random nreps=100 rearrlimit=5000000 limitperrep=yes}).

@@

\begin{figure}
\centering
   \includegraphics[width=0.9\textwidth]{RF_results_best.eps}
\caption{\small{Effect of character difference on recovering the ``normal'' topology. The y axis represents the Normalised Tree Similarity using Robinson-Fould distance for matrices with 25, 75 and 150 taxa from top to bottom respectively. The x axis represents the different character difference scenarios and tree inference method with the ``maximised'' character difference in Bayesian (red) and under maximum parsimony (orange), the ``minimised'' character difference in Bayesian (dark green) and under maximum parsimony (light green) and the ``randomised'' character difference in Bayesian (dark blue) and under maximum parsimony (light blue) for matrices of 100, 350 and 1000 characters in the panels from left to right.}}
\label{Fig:RF_results_best}
\end{figure}

\begin{figure}
\centering
   \includegraphics[width=0.9\textwidth]{Tr_results_best.eps}
\caption{Effect of character difference on recovering the ``normal'' topology. The axis are identical to figure \ref{Fig:RF_results_best} but y axis represents the Normalised Tree Similarity using Triplets distance.}
\label{Fig:Tr_results_best}
\end{figure}

@@

\begin{table}[ht]
\centering
\begin{tabular}{llr|rl}
  \hline
metric & test & bhatt.coeff & statistic & p.value \\ 
  \hline
RF & maxi:mini & 0.573 & 356436.000 & \textbf{0} \\ 
   & maxi:rand & 0.873 & 287225.000 & \textbf{0} \\ 
   & mini:rand & 0.856 & 95841.500 & \textbf{0} \\ 
  Tr & maxi:mini & 0.614 & 358800.000 & \textbf{0} \\ 
   & maxi:rand & 0.908 & 288223.500 & \textbf{0} \\ 
   & mini:rand & 0.858 & 98507.500 & \textbf{0} \\ 
   \hline
\end{tabular}
\caption{Differences between the pooled scenarios. Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correction).} 
\label{Tab_pooledscenarios_test}
\end{table}

@@

\begin{table}[ht]
\centering
\begin{tabular}{llr|rl}
  \hline
metric & test & bhatt.coeff & statistic & p.value \\ 
  \hline
RF & c100:c350 & \textbf{0.99} & 190357.500 & 1 \\ 
   & c100:c1000 & \textbf{0.98} & 174085.500 & \textbf{0.001} \\ 
   & c350:c1000 & \textbf{0.984} & 180460.000 & \textbf{0.032} \\ 
  Tr & c100:c350 & \textbf{0.961} & 166609.500 & \textbf{0} \\ 
   & c100:c1000 & \textbf{0.956} & 151389.500 & \textbf{0} \\ 
   & c350:c1000 & \textbf{0.981} & 178793.500 & \textbf{0.014} \\ 
   \hline
\end{tabular}
\caption{Differences between the pooled number of characters. Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correction).} 
\label{Tab_pooledscharacters_test}
\end{table}

@@

\begin{table}[ht]
\centering
\begin{tabular}{llr|rl}
  \hline
metric & test & bhatt.coeff & statistic & p.value \\ 
  \hline
RF & t25:t75 & \textbf{0.976} & 218421.000 & \textbf{0.012} \\ 
   & t25:t150 & \textbf{0.988} & 220529.000 & \textbf{0.004} \\ 
   & t75:t150 & \textbf{0.99} & 201037.000 & 1 \\ 
  Tr & t25:t75 & \textbf{0.976} & 233282.000 & \textbf{0} \\ 
   & t25:t150 & \textbf{0.978} & 227288.000 & \textbf{0} \\ 
   & t75:t150 & \textbf{0.992} & 194201.000 & 1 \\ 
   \hline
\end{tabular}
\caption{Differences between the pooled number of taxa. Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correction).} 
\label{Tab_pooledstaxa_test}
\end{table}

@@

\begin{table}[ht]
\centering
\begin{tabular}{llr|rl}
  \hline
metric & test & bhatt.coeff & statistic & p.value \\ 
  \hline
RF & bayesian:parsimony & 0.891 & 579437.500 & \textbf{0} \\ 
  Tr & bayesian:parsimony & \textbf{0.984} & 470621.500 & 0.084 \\ 
   \hline
\end{tabular}
\caption{Differences between the pooled methods. Bhatt.coeff is the Bhattacharrya Coefficient (probability of overlap), the statistic and the p.value are from a non-parametric wilcoxon test (with Bonferonni-Holm correction).} 
\label{Tab_pooledsmethods_test}
\end{table}

@@

\subsection{Effect of character differences on topology}
As expected, there is a significant effect of the character difference in the ability to recover the correct topology.
The character difference metric can be seen as the inverse of character correlation (see Methods): a high character difference approximates a low level of character correlation and vice versa.
When characters are correlated, one could expect the matrices to convey a strong (but potentially misleading) phylogenetic signal since every character agrees with each other and conversely, when characters are uncorrelated, one could expect them to convey a weaker phylogenetic signal with a high amount of homoplasy.
Intuitively, this would lead the ``minimised'' character difference scenario to lead to incorrect but consistent trees, the ``maximised'' scenario to lead to poorly resolved once (really homoplasic trees) and the ``randomised'' scenario to perform the best at recovering the correct topology.
Although the expected results appear to be true for a low character difference scenario, increasing the character difference surprisingly improves the ability to recover the ``normal'' topology both in terms of clade conservation ($NTS_{RF}$) and taxa placement ($NTS_{Tr}$) for both inference methods (especially in bigger matrices; Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best}).
Furthermore, the trees generated by the ``minimised'' scenario do not appear better resolved (towards any topology) than the other scenarios (see Supplementary material 3, Figs 3, 4 and 5).

@@

\subsection{Conclusion}
Correlation between characters can be induced through three main phenomena: intra-organismal relationships, selection-driven covariation or biases in coding the characters yet only the latter can be improved upon to investigate phylogenetic relationships.
Useful best practices guidelines \citep[e.g.][]{Brazeau2011,simoes2017giant} and algorithms for dealing with different types of character correlations \citep[e.g. for characters hierarchy][]{DeLaet2015,BrazeauNA} already exist.
However, with the regain of popularity in discrete morphological data and the expansion of dataset size \citep[e.g.][with more than 1000 characters each]{nithe2013,O'Leary08022013}, we can expect the correlation between characters to increase stochastically.
Moreover, because phylogenetic inference software are unable to \textit{a priori} differentiate these difference correlations, it is important to understand to what extant topologies can be induced by such bias.

>>>>

\begin{figure}[!htbp]
\centering
   \includegraphics[width=0.9\textwidth]{Figures/outline.pdf}
\caption{Outline of the simulation protocol: the first step includes both the simulation and the modification of the matrices (thin solid lines); the second step includes tree inference using MP and BPP methods (thick solid lines); the third step includes comparing the resulting tree topologies (dashed lines). $n$ and $m$ corresponds to the number of characters with a character difference $<0.25$ and $>0.75$ respectively.}
\label{Fig:outline}
\end{figure}

@@

\begin{figure}[!htbp]
\centering
   \includegraphics[width=1\textwidth]{Figures/Modif_matrix.pdf}
\caption{Example illustration of the protocol for modifying matrices. The matrices represent the pairwise character differences for 100 characters. Blue colours correspond to low character differences and orange colours correspond to high character differences. \textbf{A} - a random Birth-Death tree is simulated and used for generating the ``normal'' matrix (\textbf{B}), characters in this matrix are then removed or duplicated to favour maximised (\textbf{C}), minimised (\textbf{D}) or randomise character difference (\textbf{E}). The differences between the characters is low in \textbf{C} (minimised compared to \textbf{A}) implying a high correlation between the characters. Conversely, the character differences is high in \textbf{D} (maximised compared to \textbf{A}) implying a low correlation between the characters.}
\label{Fig:modif_matrix}
\end{figure}

\subsection{Inferring topologies}
We inferred the topologies with both BPP and MP using MrBayes \citep[v3.2.6;][]{Ronquist2012mrbayes} and PAUP* \citep[v4.0a151;][]{swofford2001paup} respectively.
For both methods, we used the arbitrarily chosen outgroup in the simulations to root our trees.
The maximum parsimony inference was run using a heuristic search with random sequence addition replicate 100 times with a limit of $5\times10^6$ rearrangements per replicates (\texttt{hsearch addseq=random nreps=100 rearrlimit=5000000 limitperrep=yes}).

@@

\begin{figure}[!htbp]
\centering
   \includegraphics[width=1\textwidth]{Figures/RF_results_best.pdf}
\caption{\small{Effect of character difference on recovering the ``normal'' topology. The y axis represents the Normalised Tree Similarity using Robinson-Fould distance for matrices with 25, 75 and 150 taxa from top to bottom respectively. The x axis represents the different character difference scenarios and tree inference method with the ``maximised'' character difference in Bayesian (red) and under maximum parsimony (orange), the ``minimised'' character difference in Bayesian (dark green) and under maximum parsimony (light green) and the ``randomised'' character difference in Bayesian (dark blue) and under maximum parsimony (light blue) for matrices of 100, 350 and 1000 characters in the panels from left to right.}}
\label{Fig:RF_results_best}
\end{figure}

\begin{figure}[!htbp]
\centering
   \includegraphics[width=1\textwidth]{Figures/Tr_results_best.pdf}
\caption{Effect of character difference on recovering the ``normal'' topology. The axis are identical to figure \ref{Fig:RF_results_best} but y axis represents the Normalised Tree Similarity using Triplets distance.}
\label{Fig:Tr_results_best}
\end{figure}

@@

\input{Tables/Tab_pooledscenarios_test.tex}

@@

\input{Tables/Tab_pooledscharacters_test.tex}

@@

\input{Tables/Tab_pooledstaxa_test.tex}

@@

\input{Tables/Tab_pooledsmethods_test.tex}

@@

\subsection{Effect of character differences on topology}
As expected, there is a significant effect of the character difference in the ability to recover the correct topology.
The character difference metric can be seen as the inverse of character correlation (see Methods \ref{CD_description}): a high character difference approximates a low level of character correlation and vice versa.
When characters are correlated, one could expect the matrices to convey a strong (but potentially misleading) phylogenetic signal since every character agrees with each other and conversely, when characters are uncorrelated, one could expect them to convey a weaker phylogenetic signal with a high amount of homoplasy.
Intuitively, this would lead the ``minimised'' character difference scenario to lead to incorrect but consistent trees, the ``maximised'' scenario to lead to poorly resolved once (really homoplasic trees) and the ``randomised'' scenario to perform the best at recovering the correct topology.
Although the expected results appear to be true for a low character difference scenario, increasing the character difference surprisingly improves the ability to recover the ``normal'' topology both in terms of clade conservation ($NTS_{RF}$) and taxa placement ($NTS_{Tr}$) for both inference methods (especially in bigger matrices; Figs \ref{Fig:RF_results_best} and \ref{Fig:Tr_results_best}).
Furthermore, the trees generated by the ``minimised'' scenario do not appear better resolved (towards any topology) than the other scenarios (see Supplementary material 3, Figs 3, 4 and 5).

@@

\subsection{Conclusion}
Correlation between characters can be induced through three main phenomena: intra-organismal relationships, selection-driven covariation or biases in coding the characters yet only the latter can be improved upon to investigate phylogenetic relationships.
Useful best practices guidelines \citep[e.g.][]{Brazeau2011,simoes2017giant} and algorithms for dealing with different types of character correlations \citep[e.g. for characters hierarchy][]{deLaet2015,BrazeauNA} already exist.
However, with the regain of popularity in discrete morphological data and the expansion of dataset size \citep[e.g.][with more than 1000 characters each]{nithe2013,O'Leary08022013}, we can expect the correlation between characters to increase stochastically.
Moreover, because phylogenetic inference software are unable to \textit{a priori} differentiate these difference correlations, it is important to understand to what extant topologies can be induced by such bias.
